### 2023-04-28 paper/case_study.tex
reorganize paper

\section{Case Studies}
\label{sec:case}

examining 3-4 cases, they are the proof that the approach works

\subsection{Reentrancy}

\subsection{Callback}

\subsection{Comparison with \texttt{modifier}}

\todo{what else?}