### 2023-05-08 paper/case_study.tex
intro

\section{Case Studies}
\label{sec:case}
examining 3-4 cases, they are the proof that the approach works

\subsection{Reentrancy}

\paragraph{Severity} talk about severity for each case

\subsection{Callback}

\subsection{Comparison with \texttt{modifier}}

\todo{what else?}