### 2023-06-03 paper/implementation.tex
misc

\section{Implementation}

We implement \lang as a preprocessor of Solidity programs annotated with
\lang's specifications (\todo{bwd ref}). Given a such input program, \lang analyzes where and
which condition needs to be inserted into the original program.
The output program can be compiled and deployed using downstream Solidity
toolchains without any further modification.

Our current implementation does not handle inlined assembly. \todo{why}

Contracts for higher-order functions is another future work when Solidity has a
proper support for lambda expressions. Without such proper support from the base language,
\todo{...}