### 2023-06-06 paper/background.tex
paper

\section{Background: The Solidity Language}

brief introduction of the Solidity language, why it is complicated

address: they are callable and can perform arbitrary computation, but we cannot always
inspect the computation other than checking the underlying numeric addressing value.

\paragraph{Challenges}
The unique features of the Solidity language and blockchain computation impose
several challenges that demand us exploring different design choices.
Here we summarize these challenges and in the next section we introduce the
design of \lang.

\begin{itemize}
  \item Rich higher-order computation:
    address are \emph{opaque first-class values carrying higher-order computation}.
  \item Distributed execution model
  \item Open-world assumption: unknown/adversarial caller, we do not know all the programs
    and cannot modify them
  \item Efficiency requirement: gas consumption
\end{itemize}