### 2023-06-25 paper/formal_model.tex
update grammar in paper

\section{Formal Model}
\label{sec:model}

the core model of the language $\lambda_\lang$

\subsection{Syntax}

\input{fig:syntax}

\subsection{Semantics}

static semantics

dynamic semantics
\yy{do we want to model the context of contracts? (instead of reasoning one contract)}

The semantics of contracts

\subsection{Compilation of Specification}

\subsection{Soundness}

what does it guarantee