### 2023-06-26 paper/formal_model.tex
update syntax

\section{Formal Model} \label{sec:model}

To articulate our approach and \lang's semantics, we present a core model
smart contract language $\lambda_\lang$ equipped with
behavioral contracts. The language modulo behavioral
contracts is entirely a subset of Solidity.
This section presents $\lambda_\lang$'s abstract syntax, static semantics, and
translation semantics.

\subsection{Syntax}

\input{fig:syntax}

\Cref{fig:syntax} shows the abtract syntax of $\lambda_\lang$.

\subsection{Static Semantics}

static semantics

dynamic semantics
\yy{do we want to model the context of contracts? (instead of reasoning one contract)}

The semantics of contracts

\subsection{Translation Semantics}

\subsection{Soundness}

what does it guarantee