### 2023-06-26 paper/case_study.tex
more formal stuff

\section{Case Studies} \label{sec:case}

We now turn our attention to real world cases and examine how \lang
can be used to consolidate smart contract programs.
These case studies are empirical evidence showing \lang is effective
in preventing attacks.

\subsection{Reentrancy}

\paragraph{Severity} talk about severity for each case

\subsection{Callback}

\subsection{Comparison with \texttt{modifier}}

\todo{what else?}