### 2023-07-09 paper/background.tex
background section [WIP]

\section{Background} 
\label{sec:background}

% \todo{brief introduction of the Solidity language, why it is complicated}

In this section, we start by briefly introducing the fundamental concepts of blockchain and the Solidity language, emphasizing their key features. 
% we first briefly introduce the background of blockchain and the Solidity language, focusing on their unique features. 
Thereafter, we discuss the challenges that arise when designing a specification and monitoring system dedicated to Solidity.

% blockchain in a nutshell (keep it short)
\paragraph{Ethereum}
% \textit{Ethereum}~\cite{buterin2014next} is a blockchain platform designed to be more flexible and programmable than its predecessor, \textit{Bitcoin}~\cite{nakamoto2008bitcoin}. 
Ethereum~\cite{buterin2014next} is a blockchain platform designed to  provide better flexibility and programmability than its precursor, Bitcoin~\cite{nakamoto2008bitcoin}.
There are two types of accounts in the Ethereum blockchain: \textit{Externally Owned Accounts} (EOAs) and \textit{Smart Contracts}. 
EOAs, controlled by private keys, usually stand for individual users. Meanwhile, smart contracts are self-running programs that enforce agreements without needing a third party. 
Actions on Ethereum are conducted through \textit{Transactions}, mostly involving interactions with smart contracts (e.g., invoking a function of a smart contract).

% Solidity in a nutshell
\paragraph{Solidity}
Solidity, a widely used programming language for smart contracts, shares syntax similarities with Java and JavaScript. 
In Solidity, a contract acts much like a class in Java, and it can contain data fields. 
The language allows functions with modifiers such as \texttt{external}, implying that a function can be invoked by an EOA or another contract, and \texttt{internal}, indicating that only other functions within the contract can call this function. 
Solidity compiler compiles a contract into bytecode.
When the bytecode is sent to the blockchain, the corresponding contract gets deployed.

% compilation --> deployment --> address --> high-order --> open world (boundry)
\paragraph{Address}
Both EOAs and Smart Contracts can be identified by a unique 160-bit address. 
the address of an EOA comes from its private key, while the one of a smart contract is decided by how the contract is deployed.

Addresses in Solidity are not just first-class values but also callable and capable of conducting complex computations. 
They play a crucial role in Solidity, enabling complex higher-order computations, flexible interactions, and efficient access control for smart contracts. 
The following example will provide an insight into how addresses are used in Solidity.

\begin{lstlisting}
interface IERC20 {
  function transfer(address to, uint amount) external;
}

contract Vault {
  address owner;

  constructor() { owner = msg.sender; }

  function deposit(address token, uint amount) {
    require(msg.sender == owner);
    IERC20(token)
      .transferFrom(msg.sender, address(this), amount);
  }
}
\end{lstlisting}

%TODO

+ as data 
+ as function parameter
+ can call to function that is not implemented in the current contract

XXX For example, a smart contract can call an external function of another smart contract using the address of the latter.
As shown above, the function \texttt{f} takes an address as an argument and calls the transfer function of the contract at the given address. 
Note that \texttt{IERC20} is an interface, which helps the compiler to generate the correct signatures for the function call.
By doing so, a smart contract can interact with other smart contracts without knowing their implementations.
It is also noteworthy that it is common to use an address of a smart contract that is deployed by others, which is beyond the control of the current contract. 
As long as the subject contract implements the interface, the current contract can call its functions.
Besides, Solidity treats addresses as first-class values, which can be passed as arguments, returned as results, stored in data fields, and so on.

% + introduce solidity in a nutshell
% + focus on address: first-class and high-order, basic access control
% + do we need to talk about modifiers?
% + gas - halting problem

% gas --> halting problem

address: they are callable and can perform arbitrary computation, but we cannot always
inspect the computation other than checking the underlying numeric addressing value.

\paragraph{Challenges}
Here we identify some major challenges imposed by the unique features of the
Solidity language and blockchain computation:

\begin{itemize}
  \item Rich higher-order computation:
    e.g., address are \emph{opaque first-class values carrying higher-order
    computation}.
  \item Distributed execution model:
    contracts are deployed on the blockchain and invoked remotely by an unknown
    or adversarial party.
  \item Open-world assumption:
    we do not have access and control of all smart contract programs.
  \item Efficiency requirement:
    deployed smart contract programs consumes ``gas'' and it is desirable
    to minimize gas consumption.
\end{itemize}

These challenges make it impossible to monitor all communication channels,
which further demands us exploring different design choices when layering
a specification and runtime monitoring system for Solidity.