### 2023-07-11 paper/case_study.tex
fix broken url

\section{Case Studies} \label{sec:case}
\begin{itemize}
	\item Overall result in preventing several real-world attacks.
\end{itemize}

We now turn our attention to real world cases and examine how \lang
can be used to consolidate smart contract programs.
These case studies are empirical evidence showing \lang is effective
in preventing attacks.

\subsection{Reentrancy}
Market.xyz readonly reentrancy attack.
Transfer return value not checked.
Li.fi attack: Arbitrary call to untrusted code (low-level call).

\paragraph{Severity} talk about severity for each case

\subsection{Callback}

\subsection{Comparison with \texttt{modifier}}

\todo{what else?}

\begin{table}
	\centering
	\begin{tabular}{cccc}
		\toprule
		\textbf{Type} & \textbf{Case} & \textbf{Date} & \textbf{Loss} \\
		\midrule
		\multirow{4}{*}{\rotatebox[origin=c]{90}{\begin{tabular}{@{}c@{}}Readonly \\Reentrancy\end{tabular}}}
		& \href{https://medium.com/coinmonks/theoretical-practical-balancer-and-read-only-reentrancy-part-1-d6a21792066c}{Sentiment} & 04/05/23 & \$1M \\
		& \href{https://twitter.com/peckshield/status/1623910257033617408} %\href{https://twitter.com/BlockSecTeam/status/1623901011680333824
		{dForce} & 02/10/23 & \$3.65M \\
		& \href{https://twitter.com/peckshield/status/1614774855999844352}{MidasCapital} & 01/16/23 & \$650K \\
		& \href{https://quillaudits.medium.com/decoding-220k-read-only-reentrancy-exploit-quillaudits-30871d728ad5}{Market.xyz} & 10/24/22 & \$220K \\

		\midrule

		\multirow{10}{*}{\rotatebox[origin=c]{90}{Arbitrary External Call}}
		& \href{https://twitter.com/hexagate_/status/1671188024607100928?cxt=HHwWgMC--e2poLEuAAAA}{MIMSpell} & 06/20/23 & \$17K \\
		& \href{https://twitter.com/HypernativeLabs/status/1633090456157401088}{Phoenix} & 03/07/23 & \$100K \\
		& \href{https://mirror.xyz/revertfinance.eth/3sdpQ3v9vEKiOjaHXUi3TdEfhleAXXlAEWeODrRHJtU}{RevertFinance} & 02/18/23 & \$30K \\
		& \href{https://twitter.com/peckshield/status/1626493024879673344}{Dexible} & 02/12/23 & \$1.5M \\
		& \href{https://twitter.com/peckshield/status/1622801412727148544}{CowSwap} & 02/07/23 & \$120K \\
		& \href{https://twitter.com/BlockSecTeam/status/1606993118901198849}{Rubic} & 12/25/22 & \$1.5M \\
		& \href{https://twitter.com/SlowMist_Team/status/1590685173477101570}{BrahTOPG} & 11/09/22 & \$89k \\
		& \href{https://twitter.com/Supremacy_CA/status/1590337718755954690}{MEV\_0ad8} & 11/08/22 & \$282k \\
		& \href{https://twitter.com/Supremacy_CA/status/1579813933669486592}{Rabby} & 10/11/22 & \$200K \\
		& \href{https://blog.li.fi/20th-march-the-exploit-e9e1c5c03eb9}{Li.Fi} & 03/20/22 & \$570K \\

		\midrule

		\multirow{5}{*}{\rotatebox[origin=c]{90}{\begin{tabular}{@{}c@{}}Return Value\\Not Check\end{tabular}}}
		& & & \\
		& & & \\
		& \href{https://medium.com/@QubitFin/protocol-exploit-report-305c34540fa3}{Qubit} & 01/28/22 & \$80M \\
		& & & \\
		& & & \\
		\bottomrule
	\end{tabular}
	\caption{Summary of case studies.\wac{The dForce link is broken. I changed it.}}
	\label{tab:case}
\end{table}