### 2023-08-11 paper/03_contract_solidity.tex

\section{\lang by Examples} \label{sec:examples}

%We now introduce \lang informally with examples.
\lang is a non-breaking extension for Solidity that additionally provides means
to specify and enforce behavioral contracts (we will often use \emph{specification} 
for behavioral contracts to avoid ambiguity).
%\subsubsection*{Design Rationale}
% With the mind of offering \xx{To offer ..} 
%To offer a better tool for Solidity programmers, 
%we have the following goals when designing \lang:
Now, we introduce the core features of \lang with examples.

\subsection{Contracts for First-Order Values}

We begin with specifying pre-conditions and post-conditions for functions
involving first-order arguments. 
Hereafter, we write \lang specifications in {\color{csspec}{\emph{italic serif style}}}, 
and ordinary Solidity program in \code{sans-serif}.
Consider the following \code{swap} example that swaps the amount \code{in} of one token to another token according to the current exchange rate. 
Callers can specify the minimal amount \code{min_out} of the output token that they expect in case the exchange rate fluctuates.

\begin{lstlisting}[language=Consol]
swap(in, min_out) returns (out)
requires in > 0
ensures out >= min_out
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}[language=Solidity]
function swap(uint in, uint min_out) returns (uint out) { ... }
\end{lstlisting}

The first three lines are \lang specifications.
The first line of the specification introduces bindings for function \code{swap}'s 
argument \spec{in}, \spec{min_out} and return value \spec{out}.
The \spec{requires}-clause specifies the pre-condition to call \code{swap},
and the \spec{ensures}-clause specifies the post-condition.
Any occurrences of \spec{in} in the conditions refer to the actual
argument value from call-sites, and similarly, occurrences of \spec{out}
refer to the actual return value.

Note that in the specification, binding names do not have to match 
those in the function definition (although in this case, they do). 
Similarly, types of arguments and returned values can often be inferred from
the definition, thus are omitted.
%\yy{types are specified in the function signature, right?}

% \subsubsection*{\textbf{Syntactic Sugars}}
\bfpara{Syntactic Sugars.}
We can liberally omit any part of the argument or return value specification,
e.g., the following specification omits the \spec{requires}-clause for \code{toWei}: %\yy{swap requires and ensures?
\begin{lstlisting}[language=Consol]
toWei(x) returns (_) ensures x < type(uint).max / 1e18
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}
function toWei(uint x) returns (uint) { ... }
\end{lstlisting}
The above specification is equivalent to the core form where the omitted 
part is simply the Boolean \code{true} expression:
\begin{lstlisting}[language=Consol]
toWei(x) returns (_) requires true ensures x < type(uint).max / 1e18
\end{lstlisting}

% \subsubsection*{\textbf{Dependent Contract}}
\bfpara{Dependent Contract.}
It is possible to write specifications where the post-condition depends on 
the arguments. 
In other words, the scope of argument bindings spans both the precondition 
and postcondition.
For example, we can write the following condition to specify monotonicity for a numeric function:
\begin{lstlisting}[language=Consol]
f(x) returns (y) ensures y > x
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}[language=Solidity]
function f(int x) returns (int) { ... }
\end{lstlisting}

%Another way is to use the `any` flat contract, which is defined as
%\code{any = { _ | true }}:
%\begin{lstlisting}
%  # {x | x < type(uint).max / 1e18} -> any
%\end{lstlisting}

% \subsubsection*{\textbf{Any Expression is Allowed}}
\bfpara{Any Expression is Allowed.}
When specifying the pre- or post-conditions, programmers are free to use
any valid Solidity expression, including but not limited
to function calls, memory operations, and built-in special variables carrying
important transaction data (e.g. \code{msg.value}) or metadata (e.g.
\code{block.timestamp}) whose values are only available at run-time.
%The programmers are free to check desired properties against with these
%transactional information.

For example, the following snippet examines the amount of Wei that is carried within \code{msg.value}
% carried 
of the transaction as the pre-condition of \code{buyTickets}:
% \yy{is the typo intended?} \yy{ensures -> requires?}\xx{revised.}
\begin{lstlisting}[language=Consol]
buyTickets(n) requires msg.value >= 1e15 * n
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}
function buyTickets(int n) payable { ... }
\end{lstlisting}
This liberty empowers programmers to check any computable properties, 
which may be beyond the capability of static verification.

%\wac{Why cannot static verification check such properties? Are we assuming that static verification only checks the violation of inline assertions?} \yy{pre-condition has different meaning in static verificaiton. I think it will not add much value to explain it. Static verification would check the post-condition assuming the pre-condition holds; in our context, pre-condition and post-condition are both enforced at runtime. Intuitively, their difference is like implication and conjunction}
%\zz{How about other special variables, such as `block.chainid` and `tx.origin`?. Perhaps we could redesign a universal syntax for these variables in Section 2.5, where we discuss where to add the message sender}
%\url{https://docs.soliditylang.org/en/develop/units-and-global-variables.html#block-and-transaction-properties}

So far, these ``flat'' contracts for first-order values are straightforward.
They provide a systematic and more readable way to write assertions wrapping around functions.

\subsection{Contracts for Higher-Order Values}\label{sec:examples-higher-order}

Now we introduce \lang's distinct feature, i.e., specifying and enforcing
rich conditions for addresses.

% \subsubsection*{\textbf{Addresses as Numeric Values}}
\bfpara{Addresses as Numeric Values.}
Resembling pointers in C/C++, addresses in Solidity are numerical values
that can be compared or computed.
Therefore, at the first approximation, specifications of addresses are 
no different from other integer values.
For instance, we can assert that the address argument value must be non-zero:
\begin{lstlisting}[language=Consol]
f(addr) requires addr != 0
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}
function f(address payable addr) { ... }
\end{lstlisting}

% \subsubsection*{\textbf{Addresses as Latent Computation}}
\bfpara{Addresses as Latent Computation.}
However, addresses in Solidity have rich higher-order behaviors: they
represent deployed external Ethereum contracts that can be invoked.
% \note{GW: need an example with return values}
Suppose we have a \code{deposit} function,
which takes an address argument and transfers money from that address. %and invokes it in the function.
With \lang, programmers can specify the condition of such latent address
calls using a \spec{where}-clause without the need of touching the function body:
\begin{lstlisting}[language=Consol]
deposit(token, amount) requires msg.sender == owner
where {
  IERC20(token).transferFrom(sender, addr, amount) returns (success)
  requires amount > 0   ensures success
}
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}
function deposit(address token, uint amount) public {
  IERC20(token).transferFrom(...); // the call is now guarded
}
\end{lstlisting}
%msg.sender, address(this), amount);
When \code{token} is considered an instance implementing the ERC20 token standard~\cite{vogelstellerEIP20TokenStandard2023},
we assert the transfer must succeed and the transferred \spec{amount} (the third argument) must be greater than zero.
We use the same \spec{requires}- and \spec{ensures}-clause syntax to specify pre-/post-conditions,
as in top-level functional specifications.
A \spec{where}-clause may contain multiple address specifications, corresponding
to multiple address arguments of a function.

Using \lang we  have decoupled specifying and enforcing of the conditions
--- there is no need to change the function body or manually insert checks
around the call, resulting in more readable and maintainable code. 
% As we will see later 
%As we will demonstrate later, in more complicated code, this separation leads to  
% more
%enhanced
%maintainability of the code and allows programmers to confidently refactor code as needed. \xx{revised}
% maintainable code and allows programmers to fearlessly refactor the code.
%\wac{Perhaps we can also mention that address spec can be declared for a storage variable?}\zz{it is a syntax sugar}GW: mentioned in sec 5

% \subsubsection*{\textbf{Single Address, Multiple Callees}}
\bfpara{Single Address, Multiple Callees.}
% \lang is expressive enough\xx{sufficiently} to specify conditions of multiple target functions
% Since the address is an instance of a contract (or a user \xx{to a user account}) containing multiple functions, programmer can call these functions at free will.
%An address is an instance of a contract (or a user account) and contains multiple functions that can be invoked by others.
% programmers \xx{by other contract? users?}.
%\xx{revised}
% can call these functions at free will.
%\note{GW: if the address is a user, do we need the interface cast?}
%\wac{We don't need since if it is an EOA, then there is no function to invoke.}
\lang is expressive to specify conditions for multiple target functions associated with a single address value.
This can be done by  specifying conditions for different callable targets
in the \spec{where}-clause. 
Consider function \code{f}'s address argument \code{addr}, the programmer can specify
conditions for \code{addr}'s functions \code{transfer} and \code{vote}:
%Other than the \code{call} method, there are other forms of low-level calls with
%different signatures, e.g., \code{delegatecall}, \code{send}, etc.
% https://docs.soliditylang.org/en/v0.8.20/units-and-global-variables.html#address-related
\begin{lstlisting}[language=Consol]
f(addr) where
{ addr.transfer(x) returns (success) requires x<=addr.balance }
{ addr.trade(msg, x) returns (success, rate) requires x>0 ensures success}
\end{lstlisting}
% \xx{filled in with different and concrete conditions}
% \xx{vote is not a good case as it requires value}
% { addr.vote(msg, y) returns (success, data) requires y!=0 ensures success}

This flexibility allows programmers to effectively 
control and enforce distinct behaviors for different targets at a finer granularity.

%\zz{I guess we should educate the audience about the difference between low-level and high-level calls at the beginning of Section 2.4. I'll take care of this later.}
%\note{GW: it seems we need to discern \code{call}s with different signatures}

% \subsubsection*{\textbf{Call Option Arguments}}
\bfpara{Call Option Arguments.}
Address calls in Solidity can 
% additionally 
take additional special arguments
such as \code{value} and \code{gas}:
\begin{lstlisting}
addr.call{value: 100, gas: 5000}(...);
\end{lstlisting}
Programmers can use \lang to specify conditions of these call option arguments 
by introducing additional bindings using the familiar syntax that is used in Solidity:
\begin{lstlisting}[language=Consol]
addr.vote{value: v, gas: g}(msg, amount) requires v>=0 && v<=addr.balance
\end{lstlisting}
% \xx{updated example with concrete conditions}
where \code{v} and \code{g} represent the actual Ether transfer value and gas value, 
whose names can be referenced in \spec{requires}/\spec{ensures}-clauses.
%Such representations simplify the specification process, enabling more 
%accurate and expressive conditions and allow more intricate specifications.

% \subsubsection*{\textbf{Low-Level Calls}}
\bfpara{Low-Level Calls.}
We have presented how \lang can be used to specify and enforce
conditions of high-level address calls, where signatures
of the callees are available.
Another form of calls in Solidity is \emph{low-level calls},
where the callee signature and arguments are encoded as raw bytes data:
\begin{lstlisting}[language=Solidity]
  bytes memory data = ...
  (bool success, bytes memory result) = addr.call(data)
\end{lstlisting}

With \lang and Solidity's \code{decode} functionality, 
% the programmer can specify conditions of low-level calls by inspecting the raw data. 
programmers can enforce the behaviors of low-level calls by attaching specifications to the raw data.%\xx{revised}
For example, by writing preconditions as a standalone function
(recall that conditions can be any expression),
we can examine the encoded signatures and arguments in a separate function 
and check it in the \spec{requires}-clause:
\begin{lstlisting}[language=Consol]
function checkPreCond(bytes memory data) returns (bool) {
  if (bytes4(data[:4]) != bytes4(ERC20SignatureData)) 
    return false; // check if data encodes ERC20 protocol
  (uint256 n, address a) = abi.decode(data[4:], (uint256, address));
  ... // check the actual arguments n and a
}
addr.call(data) requires checkPreCond(data) // precondition of addr.call
\end{lstlisting}

% \subsubsection*{\textbf{Higher-Order Functions}}
\bfpara{Higher-Order Functions.}
Likewise, functions in Solidity are first-class values, i.e., they can be used
as arguments of, or returned from, other functions.
However, 
as of Solidity 0.8.20,
% at this moment~\footnote{Solidity version 0.8.20}, 
% Solidity's 
support for first-class functions is limited to explicitly defined top-level functions.
% there is no way to write anonymous functions (lambda expressions) or nested, 
% named functions that capture variables from environments.
Writing anonymous functions (lambda expressions), nested functions, or named functions that capture variables from environments are currently not supported.
% Only explicitly defined top-level functions can be used in ``first-class'' ways.
This restriction prohibits real closure values with lexical scopes, familiarized by functional programmers.
%This limitation restricts the use of real closure values with lexical scopes, a feature that is commonly used by functional programmers.
%This restriction discourages programmers to use higher-order functions due to
%its inconvenience and verbosity.
% In principle, our treatment to addresses can be applied to functions as well.
% However, 
While our approach to addresses could be extended to functions,
due to the restriction of Solidity, higher-order functions
are rarely used in Solidity programs.
Therefore, in this paper, we focus on the contracts and monitoring of address
values, and leave the monitoring for higher-order functions as future work once
Solidity has proper support for lambda expressions~\cite{soliditydoctypes}. % \zz{cite the document mentioning Solidity is planing to support anonymous functions}\xx{cited}
%\xx{revised}
%Moreover, the restriction from the language poses challenges to seamlessly 
%implement first-class behavioral contract and monitor system that guards 
%higher-order functions (e.g. as in \cite{DBLP:conf/icfp/FindlerF02}) 
%without an expensive whole program transformation.

% \vspace{-9mm}
\subsection{\textbf{Persistent Monitoring}}
\label{sec:examples-monitoring}
\lang features persistent monitoring of address specifications,
i.e., guarded addresses are first-class citizens too -- passing
to other functions, returning from functions, or storing into mutable states, 
all preserve the attached pre-/post-conditions.
These conditions will be checked whenever the address is called, even remotely 
from where the conditions were attached.

%For instance,
%these conditions will be enforced when the address has been passed to other functions, returned from the current function, or stored in memory. \xx{revised}
% the address has been passed to other functions, returned from the current function, or stored in storage.

%\lang tracks the use of guarded address arguments
%even if they are passed down to other functions which potentially call the address.
% For instance, 
% \subsubsection*{\textbf{Passing Guarded Addresses}}
\bfpara{Passing Guarded Addresses.}
In the following example, the programmer has attached specifications
to the public function \code{deposit} but not to the \code{actualDeposit}.
The private function \code{actualDeposit} consumes the address passed
from \code{deposit} and invokes functions on the address.
\begin{lstlisting}[language=Consol]
deposit(token, amount) 
where { IERC20(token).transferTo(addr, amt) requires amt > 10 }
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}[language=Solidity]
function deposit(address token, uint amount) public {
  address token2 = token;
  actualDeposit(token2, amount - 10);
}
function actualDeposit(address token2, uint amount) private {
  IERC20(token2).transferTo(.., amount); // call happens here
}
\end{lstlisting}
% \xx{naming: foo, and two tokens are confusing}
% \xx{added a concrete condition. move comments to the text? ->}
%\yy{I was expection something like \spec{amount>=0}} 
Although the condition for address \code{token} is specified for \code{deposit},
the call of \code{transferTo} happens remotely in \code{actualDeposit},
via indirect value flows.
\lang ensures the precondition of the address call (\spec{amt >= 0}) 
attached from  \code{deposit} is preserved and checked in \code{actualDeposit}.
%\lang guarantees this by dynamically tracking the value flow.

It is possible that \code{actualDeposit} is invoked by other functions. In such cases,
different conditions may be checked, depending on the actual address argument \code{token2}.

% \subsubsection*{\textbf{Returning Guarded Addresses}}
\bfpara{Returning Guarded Addresses.}
%Addresses can be returned functions as well. 
There are two ways to return a guarded address value. First,
returning a guarded address whose specification was attached at a previous point,
e.g., by other functions or at the time when the address is used as an argument.
Consider this identity function of addresses, which requires its address argument
to be monotonic when called: %\xx{revised}
\begin{lstlisting}[language=Solidity]
interface IMono { function f(int) returns (int); }
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}[language=Consol]
id(addr1) returns (addr2) 
where { IMono(addr1).f(x) returns (y) requires y >= x }
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}[language=Solidity]
function id(address addr1) returns (address) { return addr1; }
\end{lstlisting}
Then, when \code{addr1} is returned to the caller of \code{id}, the same condition is still attached.
At a later point, when the received address is invoked, the monotonic check
will be enforced.
% accompanied\xx{attached?}.
%For example, when the \code{addr1} (identical to \code{addr1}) is returned 
%from the function \code{id}, the specifications initially attached to \code{addr1}
%persist with \code{addr2}. 
% \xx{return from where, still attached when invoked by xxx }

The second way is to directly attach conditions to the returned address.
Using the same identify function as an example, we can specify monotonicity 
of the returned address \spec{addr2} (instead of \spec{addr1}):
\begin{lstlisting}[language=Consol]
id(addr1) returns (addr2) where { I(addr2).f(x) returns (y) requires y>=x }
\end{lstlisting}
Similarly, at a later point, when the received address is invoked, the monotonic check
will be enforced:
\begin{lstlisting}[language=Solidity]
address c = id(...)
int y = IMono(c).f(x) // checks condition y >= x
\end{lstlisting}
Moreover, it is possible to attach different precondition and postcondition
to the same address value.
%\yy{changed low-level call to high-level call}

%\note{GW: what is the first time we talk about storage?}

Allowing attaching persistent specifications to return addresses is a powerful notion.
%\todo{why it is useful?}\wac{Added}
Addresses may flow through the business logic of contracts. 
\lang liberates developers from meticulously tracking
address flows and verbosely checking the
critical conditions at every address call site. 
Consequently, developers can focus on the overall business logic,
delivering code with decent readability and maintainability (see case studies \Cref{sec:case}).
%\wac{Refer to a case study maybe.}\note{GW: yes}

Additionally, a guarded address can be stored in storage,
retrieved, and called later. When the call happens, \lang preserves the checks 
as they are remotely specified.

% \subsubsection*{\textbf{Extent of Effective Monitoring}}
\bfpara{Extent of Effective Monitoring.}
In \Cref{sec:translation}, we discuss how \lang implements persistent address
specification monitoring using a whole-program transformation.
Nevertheless, there are cases where tracking specifications attached to addresses
becomes impossible in a distributed setting, e.g., when a guarded address value 
is passed to another external contract program that is unknown to \lang.
In such cases, \lang cannot  monitor how the external contract program uses the address.
%\xx{revised}
However, programmers can still trust \lang within the scope of the annotated
contract program. \lang ensures effective and persistent monitoring of address specifications
as long as the address calls happen within the current contract.
In \Cref{sec:correctness}, we depict the boundary of effective monitoring after 
explaining the translation semantics.
%\yy{Can we justify it is often sufficient to focus on the contract program under development? Refer back to the "open-world" assumption?}
%\lang's implementation maintains the same interface of the contract to observers.

\iffalse
Nevertheless, such checks against \code{addr} cannot be performed when \code{f}
is called, even though \code{addr} is a proper argument of \code{f}.
Because \code{addr} in this use case is effectively a function, and
it is undecidable to statically examine its properties in general.
Additionally, we may not have access to the source code of the contract
represented by \code{addr}.
\begin{lstlisting}[language=Consol]
f(addr) where
{ addr.call(msg, amount) returns (flag, data)
  requires amount > 100
  ensures flag == true }
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}
function f(address payable addr) {
  ... // omitting statements before the call
  (bool success, bytes memory data) =
    addr.call(abi.encodeWithSignature(
        "foo(string, uint256)", "call foo", 128
      )
    );
  ... // omitting statements after the call
}
\end{lstlisting}
\lang addresses this issue by adopting lazy checks
\cite{DBLP:conf/icfp/FindlerF02} that only take place
when the address is actually invoked \note{GW: fwd ref to impl}.
The above snippet specifies the condition of \code{addr}
in the \code{where}-clause, which requires that
the \code{amount} argument must be greater than 100 and
the address call must succeed.
\fi

\iffalse
It would be convenient to directly enforces the success of the call and omit
the data:
\begin{lstlisting}
  # { a | { {value, gas}(arg) | <pre-cond> } -> { (true, _) | true } }
  function f(address a) { ... }
\end{lstlisting}

\zz{just FYI, `a.f` can also specify `value`, `gas`, etc.}

\begin{lstlisting}
  # { a.f | pre-cond -> post-cond }
  # { a.g | pre-cond -> post-cond }
  function f(address a) { ... }
\end{lstlisting}

\wac{We may need an example here since the necessity to designate the callee function may not be obvious to reviewers.}
\fi

\iffalse
\paragraph{Second-Class Monitoring}
%\note{GW: an easy option is to discard enforcement of escaped addresses at all.
%But it seems we can guard escaped addresses within the current contract scope.
%We need to think about layered/stacked contracts of addresses (lax/picky).}
Addresses, as first-class values, can of course escape from a function, e.g.
by returning from the function or storing in external data structures.
However, it is not always possible to enforce the specification
of addresses when escaped to third-parties.
For example, consider a function that returns an address equipped with
conditions:
\begin{lstlisting}
EXAMPLE TODO
function f(address a) returns (address) { return a; }
\end{lstlisting}
In this case, if the return address value crosses the boundary to other
Ethereum programs (by external calls initiated by a third-party) that we do not
have control, we are impotence to monitor if the conditions are violated.

In spite of that, \lang does guarantee that addresses are completely monitored
within the current Ethereum program's calling context, following a stack
discipline.
In other words, for an argument addresses \code{addr} of function \code{f},
\lang is able to detect any violation of \code{addr}'s conditions before
\code{f} returns (even if the call happens in other functions).
We called this mechanism \emph{second-class monitoring}, borrowing Strachey's taxonomy
of first- and second-class objects \cite{DBLP:journals/lisp/Strachey00}.
This is already useful to prevent \note{XXX} \note{fwd ref}, especially when
the address comes from an untrusted party.

Similarly, when a guarded address is used as an argument to another address
call, \note{we lose the control of ...}
\fi

\iffalse
	\begin{lstlisting}
function map(uint[] memory data, function (uint) pure returns (uint) f)
  internal pure returns (uint[] memory r)
{
  r = new uint[](data.length);
  for (uint i = 0; i < data.length; i++) {
    r[i] = f(data[i]);
  }
}
\end{lstlisting}

We can specify the contract for function arguments too:
\begin{lstlisting}
  # { f | {x | x < 0} -> {y | y > 0} }
  function map(int[] memory data, function (int) pure returns (int) f) { ... }
\end{lstlisting}
It might be too verbose -- so we can define those predicates separately for better readability/maintainability:
\begin{lstlisting}
  function greaterThanZero(int x) pure returns (bool) {
    return x > 0;
  }
  function smallerThanZero(int x) pure returns (bool) {
    return x < 0;
  }
  # { f | smallerThanZero -> greaterThanZero }
  function map(int[] memory data, function (int) pure returns (int) f) { ... }
\end{lstlisting}

Functions contracts can be higher-order --
it can take other function contracts
as part of the spec. For example
\zz{coooool!}
\begin{lstlisting}
  # TODO
  function f(function (function (int) returns (int) g) h) { ... }
\end{lstlisting}

Function contracts are first-class -- so if this guarded function is escaped
(e.g. by returning), the contract is still enforced:
\begin{lstlisting}
NEED AN EXAMPLE
\end{lstlisting}
\fi

\iffalse
\subsection{Comparison with \texttt{modifier}}

\lang is a significant extension of Solidity which has different use scenarios from \texttt{modifier}s.
Solidity \texttt{modifier}s are designed to change the behaviors of functions in a declarative way~\cite{soliditymod}, while \lang is introduced to declare specifications (pre/postconditions) to function invocations and its higher order arguments and return values.
\lang differs from \texttt{modifier}s in the following aspects:
First, \lang allows developers to specify postconditions for an invocation of a function.
Second, \lang can support declaring specifications on higher-order values.
Third, \lang features persistent monitoring of higher-order value specifications.
The specifications are attached to higher-order values instead of only being checked at the declared location.
This enables consistent specification enforcement throughout the entire contract.
Such features are not supported by \texttt{modifier}s and are unique to \lang, offering developers more powerful means to declare high-level behavior specifications.
\fi

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End: