### 2023-08-11 paper/05_implementation.tex

\section{Implementation and Optimizations}\label{sec:impl}

We implement \lang as a preprocessor of Solidity programs annotated with
\lang's specifications. Given an input program, \lang generates a Solidity
program following the translation outlined in \Cref{sec:model}.
The output programs can be compiled and deployed using off-the-shelf Solidity
toolchains without any further modification.
\lang is implemented with \code{solc-typed-ast}~\cite{solctypedast},
% the help of \note{X} parser, 
which can also reify a modified AST back to a Solidity program.
%\xx{the \lang spec is parsed with~\cite{antlr4}}

Our implementation handles Solidity features that are not covered by
the formalization, e.g., mappings, structs, and low-level calls.
Additionally, programmers can attach specifications to storage fields,
which can be desugared to core forms. This is useful to produce guarded
addresses at contract initialization time.

%\todo{syntax sugar: top-level initialization spec --> translate constant to immutable --> add function}

%\todo{what we didn't model in \Cref{sec:model}: mapping, structs, arrays, low-level calls, HO functions}
%Our current implementation does not handle inlined assembly. \todo{why}

%Contracts for higher-order functions are another future work when Solidity has 
%proper support for lambda expressions. Without such proper support from the base language,
%\todo{...}

% \subsubsection*{\textbf{Optimizations}} 
% \smallskip \noindent \textbf{Optimizations.}
\bfpara{Optimizations.}
The translation presented in \Cref{fig:translation} works with a set of runtime functions.
These functions cooperate with the runtime representation of guarded addresses.
We now discuss an optimized implementation that \emph{does not incur additional storage overhead}
in the generated code.

The optimization exploits the fact that the smallest unit for storage is 256 bits 
in the Ethereum Virtual Machine. Storing or loading for data smaller than 256 bits
incurs the same gas overhead as data of 256 bits.
However, the size of an address is 160 bits, which spares additional 96 bits that
can be used to encode specifications up to 96 predicates.
Therefore, we use \code{uint256} to represent guarded addresses, where the 96 MSB
encodes attached specifications and 160 LSB is preserved for addresses.

With this representation, our translation assigns numbers to conditions of addresses.
The correspondence between the assigned number and their runtime functions is emitted 
in the generated code.
Those runtime functions then can be easily implemented: The $\mathit{wrap}$ function coerces a 
\code{uint160} to \code{uint256}, and $\mathit{unwrap}$ function discards the 96 MSB of an \code{uint256} value.
The $\mathit{attachSpec}$ function only modifies the 96 MSB of a guarded address representation.
The $\mathit{dispatch}$ function can check which bits in the 96 MSB of the guarded addresses
are set, therefore knows which predicates to check.

The total number of conditions attached to addresses cannot exceed 96.
However, so far, we have not encountered real cases that need to use more than 96 address conditions. 
If with more conditions, our translation could fall back to the naive encoding discussed in \Cref{sec:translation}.

\iffalse
+ representing guarded address
+ eliminate redundant calls - the compiler should do so ... (e.g., from guarded to worker)
+ combining pre-/post-conditions in different functions (many functions share the same precondition), 
  to reduce the number of specs 
+ Syntax sugar?
\fi
%Question: where to place modifier?