### 2023-08-11 paper/fig_sturby_intro.tex

% IChainlinkAggregator private constant STETH  = IChainlinkAggregator(0x86392d...);

% ${\color{csspec} \mathit{STETH.latestRoundData()} }$
% ${\color{csspec} \mathit{returns (roundId, answer, startedAt, updatedAt, answeredInRound)} }$
% ${\color{csspec} \mathit{ensures (updatedAt > block.timestamp - 1 days) \&\& (answer > 0)} }$
% IChainlinkAggregator private constant STETH 
%   = IChainlinkAggregator(0x86392d...);
\begin{figure*}[t]
\begin{subfigure}[t]{0.49\textwidth}
\begin{lstlisting}[numbers=left,stepnumber=1,xleftmargin=1em,numberstyle=\ttfamily\color{gray},numbersep=3pt]
function getPrice(address chainklink) returns (uint256) {
  (_, uint256 ethPrice, _, uint256 updatedAt, _) = 
    IChainlinkAggregator(chainlink).latestRoundData();
  require(updatedAt > block.timestamp - 1 days);
  require(basePrice > 0);

  uint256 price = ORACLE.getRate() * ethPrice;

  $\HLBox{\texttt{\textbf{require}(price * 95 / 100 < ORACLE.getLatestPrice()}}$
  $\HLBox{\texttt{\ \  \&\& price * 105 / 100 > ORACLE.getLatestPrice());}}$

  return price;
}
\end{lstlisting}
\caption{Source code of readonly reentrancy vulnerability, and its fix (lines 10-11) 
with \code{require}, ensuring the price fluctuation within 5\%.}
\label{fig:sturby_buggy}
\end{subfigure}%
\hfill
\begin{subfigure}[t]{0.49\textwidth}

\begin{lstlisting}[language=Consol,numbers=left,stepnumber=1,xleftmargin=0.8em,numberstyle=\ttfamily\color{gray},numbersep=3pt,firstnumber=1]
getPrice(chainlink) returns (price)
ensures price * 95 / 100 < ORACLE.getLatestPrice() &&
       price * 105 / 100 > ORACLE.getLatestPrice()
where {
  IChainlinkAggregator(chainlink).latestRoundData()
  returns (_, answer, _, updatedAt, _)
  ensures updatedAt > block.timestamp - 1 days && answer > 0
}
\end{lstlisting}
\vspace{-0.25em}
\begin{lstlisting}[numbers=left,stepnumber=1,xleftmargin=1em,numberstyle=\ttfamily\color{gray},numbersep=3pt,firstnumber=9]
function getPrice(address chainlink) returns (uint256) {
  (, uint256 ethPrice, , , ) = 
    IChainlinkAggregator(chainlink).latestRoundData();
  return ORACLE.getRate() * ethPrice;
}
\end{lstlisting}
\caption{The fix with \lang specifications (lines 1--8), decoupling the specification and business logic.}
\label{fig:sturby_fix}
\end{subfigure}
\caption{Comparison of fixes for Sturdy (simplified) source code with assertions 
and \lang specification with enhanced readability and maintainability (see \Cref{sec:case} for detailed analysis).}
\label{fig:sturbyIntro}
\end{figure*}