### 2023-10-29 paper-icse24/09_conclusion.tex

\section{Conclusion}

In this paper, we propose a specification and monitoring system, \lang,
for the Solidity smart contract programming language.
\lang supports attaching and enforcing specifications for both
top-level functions and address values.
\lang persistently monitors address calls via a whole-program
transformation, which ensures any violation of address call conditions
in the current contract scope is captured.
We examine the effectiveness and gas efficiency using 10 real-world attacks.
By replaying existing attack transactions, \lang-patched contracts successfully
defend the attack with only 0.98\% more gas consumption.

% vs FanLong 0.43%
% ERC20/ERC721/ERC1202 invaranits 0.43\%