### 2023-10-29 paper-icse24/fig_dexible.tex

\begin{figure}[t]
\begin{lstlisting}[numbers=left,stepnumber=1,xleftmargin=1em,numberstyle=\ttfamily\color{gray},numbersep=3pt]
function fill(SwapTypes.SwapRequest calldata request, 
SwapMeta memory meta) external onlySelf returns (SwapMeta memory)  {
  //some checks
  for(uint i=0;i<request.routes.length;++i) {
    SwapTypes.RouterRequest calldata rr = request.routes[i];
    IERC20(rr.routeAmount.token).safeApprove
        (rr.spender, rr.routeAmount.amount);
    (bool s, ) = rr.router.call(rr.routerData);  // vulnerable
    ...
    }
  ...
  require(meta.outAmount >= request.tokenOut.amount,
  "Insufficient output generated");
}
\end{lstlisting}

\caption{A Simplified Demonstration of an Arbitrary External Call Bug: Dexible Contracts}
\label{fig:dexibleHack}
\end{figure}

\begin{comment}
\begin{figure}[t]
\begin{lstlisting}[numbers=left,stepnumber=1,xleftmargin=1em,numberstyle=\ttfamily\color{gray},numbersep=3pt]
struct SwapData {
  address callTo;
  bytes callData;
  ...
}
function swap(bytes32 transactionId, SwapData calldata _swap) 
  internal {
  ... // some balance check
  meta.outAmount = request.tokenOut.token.balanceOf(address(this));
  (bool success, bytes memory res) = _swap.callTo.call{
    value: nativeValue
  }(_swap.callData);
  if (!success) {
    string memory reason = LibUtil.getRevertMsg(res);
    revert(reason);
  }
  ...
}

\end{lstlisting}

\caption{A Simplified Demonstration of an Arbitrary External Call Bug: LiFi Contracts}
\label{fig:lifi_example}
\end{figure}

\begin{figure}[t]
\label{fig:lifi_solution}
\begin{lstlisting}[language=Consol]
swap (transactionId, _swap) 
requires {(_swap.callTo == XXX && _swap.callData[:4] == YYY) || ...}
\end{lstlisting}

\begin{lstlisting}[language=Consol]
swap (transactionId, _swap) 
where { _swap.callTo.call(data) returns (success, res)
    requires {_swap.callTo == XXX && data[:4] == YYY) || ...}
    ensure {success}
}
\end{lstlisting}
\caption{\lang Specifications for LiFi attack.}
\label{fig:lifi_solution}
\end{figure}

\end{comment}