### 2023-10-29 paper-icse24/fig_sturby.tex

\begin{figure}[t]
\begin{lstlisting}[numbers=left,stepnumber=1,xleftmargin=1em,numberstyle=\ttfamily\color{gray},numbersep=3pt]
function getRate() public view returns (uint256) {
  uint256 _poolTokenBalance = _getPoolTokenBalance(); // not uptated
  return _calculateRate(_poolTokenBalance);
}

function exitPool(address sender, address recipient, 
  uint256 amount) nonReentrant external {
  (bool success, ) = recipient.call{value: amount}(""); // reentrancy
  _updatePoolTokenBalance(amount);
}
\end{lstlisting}

\caption{A Simplified Demonstration of an Readonly Reentrancy: Balancer Contracts}
\label{fig:sturbyHack}
\end{figure}